Verifying Hierarchical Cache Coherence Protocols: a theorem-prover based modular, reusable proof

### 2014-02-05 paper/Deadlock.tex
intuition

In Section \ref{safety} we showed how the MSI protocol of Figure \ref{trans}
obeys the Store atomicity theorem. In addition to this criteria, we must
establish that the system makes forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where none of the transitions of Figure
\ref{trans} apply or
\item Strong forward progress: At any time, eventually a new request from some processor will be
processed
\end{enumerate}

In order to prove strong forward progress, we need to
establish some more conditions on the strategy to execute the transitions of
Figure \ref{trans} in each cache -- the caches must ensure local
starvation-freedom in the processing of messages when sufficient resources are
available. We do not discuss this issue in this paper and instead focus on 
weak forward progress.

@@

This condition ensures that every request gets a response eventually. Using
this invariant, it is easy to see that there will always be some cache in the
system which can either send a request to another cache (ChildSendReq or
ParentSendReq) or process a request (ParentRecvReq, ChildRecvReq or
ChildDropReq) or accept a response from another cache (ChildRecvResp or
ParentRecvResp). We do not give the formal proof here due to lack of space.

>>>>

In Section~\ref{safety} we showed how the MSI protocol of Figure~\ref{trans}
obeys the store atomicity theorem. To complete the picture, we must also show
that it avoids deadlock, \ie{} the protocol makes forward progress.

Forward progress can be either of the following:
\begin{enumerate}
\item Weak forward progress: The system never reaches a state where
  none of the transitions of Figure~\ref{trans} apply, or
\item Strong forward progress: At any time, eventually a new request
  from some processor will be processed.
\end{enumerate}

To prove strong forward progress, we need some guarantees on the choice of
executions of the transitions Figure~\ref{trans} in each cache. Each cache must
ensure that it does not avoid receiving a particular available request from one
of its children indefinitely, when the guard for the corresponding transition
(ParentRecvReq) is true. Besides, all the transitions in Figure \ref{trans} are
voluntary in nature. Thus, the system can execute a sequence of transitions
which simply moves, say, the \Mo{} permission among the children of a cache,
without doing any further work with that permission
%This can potentially be enforced via during hardware
%compilation;
For this paper we focus only on weak forward progress, as we have not modeled
the algorithm for correctly choosing the transitions to guarantee strong
forward progress in Coq.

@@

This condition ensures that every request gets a response eventually. Using
this invariant, it is easy to see that there will always be some cache in the
system which can either send a request to another cache if the cache is not
waiting for any response from the latter cache (ChildSendReq or ParentSendReq),
or accept a response from another cache if such a response is available
(ChildRecvResp or ParentRecvResp). But if all the caches are waiting for a
response, and no response is available, then the leaf caches can execute
ChildRecvReq transition (as the leaf does not have any children, and this
transition is executed even if a cache is waiting for a response from its
parent). Therefore, some transition can always be executed, proving weak
forward progress.
